---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%"
)
```

# SETrNCBN

<!-- badges: start -->

<!-- badges: end -->

The goal of SETrNCBN is to simplify calculations and make graphs for QA/QC and communication of NCBN Surface Elevation Table (SET) data.

This package is under development [GitHub, here](https://github.com/laura-feher/NCBN_SETr).

This README still needs a lot of attention too; the example below is very minimal.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("laura-feher/NCBN_SETr")
```

## Example

This is a basic example which shows you how to make a simple graph of change since the first reading at each SET:

```{r example, message = FALSE, warning = FALSE}
library(SETrNCBN)

# first, perform cumulative change calculations
cumu_set <- calc_change_cumu(example_sets)

# now plot cumulative change by SET
plot_cumu_set(example_sets)

# or by arm, at a single SET
plot_cumu_arm(example_sets)
```
